---
title: "Looking at tweets by @year_progress 2019-2020"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
stub: "year-progress-tweets-viz"
output:
  github_document:
    html_preview: true
    toc: false
    fig_width: 5
    fig_height: 5
    md_extensions: +emoji+bracketed_spans+inline_notes
keywords: [lubridate, ggplot, twitter, api, dataviz]
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)


library(dplyr)

```

<!-- badges: start -->
<!-- badges: end -->


I was randomly interested in whether there might be patterns in the amount of interaction with tweets from the **@year_progress** bot during 2020 :scream:,
and what the shape of the year might look like compared to 2019.


This gave me a chance to review the current landscape of `R` clients for the twitter API :disappointed:.
After doing this review, I decided it would be easier and more satisfying for me to script my own access to [v2 of the API][tw-api].

[tw-api]: https://developer.twitter.com/en/docs/twitter-api/early-access


This meant more practice with `httr`, error handling, working with lists with the frankly brilliant [`purrr` toolkit][purrr], and with writing functions.
Happy as a :pig: in :poop:, _tbqhwy_.


It's also chance for me to try various other things I haven't really used before:

* Creating a GitHub README as an `.Rmd` file (this very file here),
  instead of my usual `.md`, so I might incorporate some output charts
* Using a couple of [Pandoc's Markdown extensions][pdmd] ^[1][1]^
* GitHub Actions (set up using the `usethis` helper function)
* Making _a whole actual package_ :package:, just for a little bit of fun like
  this
* Learning a bit more about testing with `testthat`
* Yet more practice at package architecture and requirements,
  and appeasing R CMD check
* Doing some hopefully nice [dataviz]{.sparkle} :bar_chart: with `ggplot2`, which   I still don't feel like I have really used _well_ yet.
  My charts usually look way too crappy for my liking.


[1]: # "We've got md_extensions: +emoji+bracketed_spans+inline_notes going on in the yaml here"
[purrr]: https://purrr.tidyverse.org/
[pdmd]: https://pandoc.org/MANUAL.html#pandocs-markdown


If you suspect that this whole idea was essentially driven by pre-Christmas/end-of-term task avoidance, angst and procrastination,
you would be correct.



```{r, get-dtf, eval = FALSE}
# Set variables and run overall query -------------------------------------


# Obtained originally by:
# user_id <- get_user_id(username = "year_progress")
# It isn't going to change, so once obtained just hardcode it:
year_progress_id <- "3233484298"


start_2019 <- "2019-01-01T00:00:00Z"

# This doesn't work with the Twitter API
# https://github.com/tidyverse/lubridate/issues/941
# 
# start_2019 <- lubridate::make_datetime(2019) %>% 
#   lubridate::format_ISO8601(usetz = TRUE)


# see `R/get_user_twetes.R` for how this works
dtf <- get_user_twetes(user_id = year_progress_id, start = start_2019) %>%
  compact() %>%
  unpack_list()       # <- in `R/helpers.R`
```



```{r, load-dtf, include = FALSE}
# myrmidon::save_it(dtf)
dtf <- readRDS(here::here("rds_data", "dtf.Rds"))
```

I realised the other day that it would be fine to call a dataframe `dtf` instead of boring old `df` :smirk:...

```{r, dtf-tweet, echo = FALSE}
tweetrmd::include_tweet("https://twitter.com/ludictech/status/1338815874770329602?s=20")
```

```{r, glimpse-dtf}
glimpse(dtf)

```


So let's lick this dataframe into shape so it's ready for visualising.





.

.

.

.



You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.
